---
title: "Exploring Workforce Dynamics: An HR Data Study"
output: html_document
date: "2023-11-11"
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
```

#### Here we are going to perform analysis on hr dataset, which help the organisation to understand why the employees are leaving the company and how well employees are performing so that they can make improvement in their business strategy.

### 1. KPI (Key Performing Indicator)
Firstly, we will be looking the basic overview of a company as we can see total 207 employees are working out of 311, which we can say the attrition rate 34.11% in this organsation.

``` {r include = FALSE}
design <- function() {
  color_list <- list(
    link_water = "#D4E6F1",
    biscay = "#154360",
    gossip = "#DAF7A6",
    amber = "#FFC300",
    outrageous_orange = "#FF5733",
    monza = "#C70039"
  )
  
  font_list = list(
    text = "monospaced",
    title = "arial"
  )
  
  element_of = list(
    color = color_list,
    font = font_list,
    data = "C:\\Users\\HP\\OneDrive\\Google Data Analytics\\Project\\HR Analysis\\dataset\\HR_Dataset.csv"
  )
  
  return(element_of)
}
```

``` {r message = FALSE, include = FALSE}
hr_data = read.csv(design()$data)

kpi <- hr_data %>%
  summarise(Employee_Count = length(Employee_Name), 
            Attrition_Count = sum(AttritionCount),
            Attrition_Rate = round(((Attrition_Count / Employee_Count) * 100), digits = 2),
            Active_Employee = (Employee_Count - Attrition_Count))

kpi_measures <- data.frame(Names = c("Employee Count", "Attrition Count", 
                                     "Attrition Rate", "Active Employee"), 
                          Values = c(1), 
                          Counts = c(kpi$Employee_Count, kpi$Attrition_Count, 
                                     kpi$Attrition_Rate, kpi$Active_Employee))
```

```{r kpi_measures, echo=FALSE, warning=FALSE}
ggplot(kpi_measures, aes(x = Values, y = Names)) +
  geom_bar(stat = "identity", fill = "white") +  # Fixed length bar with a width of 0.5
  geom_text(aes(label = Counts), hjust = 0.9, size = 11, 
            family = design()$font$title, color = design()$color$biscay,
            fontface = "bold") + 
  scale_x_continuous(breaks = NULL) + 
  theme_minimal() + 
  theme (
    axis.text.y.left = element_text(family = design()$font$text, 
                                    size = 15, hjust = 0, 
                                    color = design()$color$amber), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank()
  )
```

### 2. Attrition by Gender:
Next, we saw more female is leaving the company than male which is almost 20% in higher side.

``` {r message = FALSE, include = FALSE}
att_gender <- hr_data %>%
  group_by(Sex) %>%
  summarise(Attrition_Count = sum(AttritionCount))
```

``` {r att_gender, echo = FALSE, warning=FALSE}
ggplot(data = att_gender, aes(x = Attrition_Count, y = Sex)) + 
  geom_segment(aes(xend = 0, yend = Sex), color = design()$color$gossip, linewidth = 4) + 
  geom_point(color = design()$color$amber, size = 15) + 
  geom_text(aes(label = Attrition_Count), color = design()$color$biscay) + 
  labs (
    title = "Attrition by Gender",
    subtitle = "Numbers of empployee leave the organisation by different gender.",
    x = "Attrition Count",
    y = "Sex"
  ) + 
  theme_minimal() + 
  theme (
    plot.title = element_text(family = design()$font$title, size = 20, 
                              color = design()$color$biscay, face = "bold"),
    plot.subtitle = element_text(family = design()$font$text, size = 15, 
                                 color = design()$color$biscay),
    axis.text = element_text(family = design()$font$text, size = 15, 
                             color = design()$color$biscay, face = "bold"),
    axis.title = element_text(family = design()$font$text, size = 17, 
                              color = design()$color$biscay),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.x = element_blank()
  )
```

### 3. Best Recruting Source
Here, we will be looking the numerous source of hiring is being done by the company in which LinkedIn and Indeed is on higher side which is showing a best prefrence of employee.

``` {r include = FALSE}
hiring_source <- hr_data %>%
  group_by(RecruitmentSource) %>%
  summarise(Employee_Hire = length(Employee_Name)) %>%
  arrange(desc(Employee_Hire))
```

``` {r hiring_source, echo = FALSE, warning=FALSE}
ggplot(data = hiring_source, aes(x = Employee_Hire, y = RecruitmentSource)) + 
  geom_bar(stat = "identity", fill = design()$color$amber, width = 0.7) + 
  geom_text(aes(label = Employee_Hire), nudge_x = 5, size = 4, 
            color = design()$color$biscay, fontface = "bold") + 
  labs(
    title = "Employees Hired / Various Recruitment Sources",
    x = "Employees Hired",
    y = "Recruitment Sources"
  ) + 
  theme_minimal() + 
  theme(
    plot.title = element_text(family = design()$font$title, size = 20, hjust = 0.5, 
                              color = design()$color$biscay, face = "bold"),
    axis.text = element_text(family = design()$font$text, size = 10, 
                             color = design()$color$biscay, face = "bold"),
    axis.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.x = element_blank()
  )
```


### 4. Company Diversion
Next, we will be exploring how diverse company which showing people are working here from different age group in which majory of higher working employee is under age of 50 which is one-third of the total employees.

```{r include=FALSE, message=FALSE}
# Creating Age Group Extra Column in our HR_Dataset . . .
data <- hr_data %>%
  mutate(AgeGroup = case_when(
    Age > 30 & Age <= 40 ~"31-40",
    Age > 40 & Age <= 50 ~"41-50",
    Age > 50 & Age <= 60 ~ "51-60",
    TRUE ~ "Over 60"
  ))

# Extracting Insights . . .
comp_diversion <- data %>%
  group_by(AgeGroup, Sex) %>%
  summarise(Employee_Count = n())
```

```{r comp_diversion, echo = FALSE, warning=FALSE}
ggplot(data = comp_diversion, aes(x = "", y = Employee_Count, fill = Sex)) +
  geom_bar(stat = "identity", position = "fill", width = 0.8, 
           color = design()$color$link_water) + 
  geom_text(aes(label = Employee_Count), position = position_fill(vjust = 0.5), 
            size = 4, color = design()$color$link_water) +
  coord_polar(theta = "y") +
  facet_wrap(~AgeGroup) +
  labs(
    title = "Company Diversity",
    subtitle = "Employee Count by Various Age Groups",
    fill = "Sex",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() + 
  theme (
    plot.title = element_text(family = design()$font$title, size = 20, hjust = 0, 
                              color = design()$color$biscay, face = "bold"),
    plot.subtitle = element_text(family = design()$font$text, size = 15, 
                                 color = design()$color$biscay),
    axis.text = element_text(family = design()$font$text, size = 10, 
                             color = design()$color$biscay, face = "bold"),
    axis.text.x = element_blank(),
    panel.grid = element_blank()
  )
```

### 5. Work Life Balance
It is really important to be more productive in a company achieving this you must have good work life balance in your life. By looking at the data we can see those employees took leaves are showing more productive in the company as they did more project.

```{r include=FALSE}
wlb <- hr_data %>%
  group_by(PerformanceScore) %>%
  summarise(Absences = sum(Absences),
            Emp_Satisfaction = sum(EmpSatisfaction),
            Sp_Project = sum(SpecialProjectsCount)) %>%
  arrange(desc(Absences))

wlb_cluster = tidyr::pivot_longer(wlb, cols = c(Absences, Emp_Satisfaction, Sp_Project), 
                                  names_to = "Variable", values_to = "Value")
```

```{r wlb_cluster, echo=FALSE, warning=FALSE}
ggplot(data = wlb_cluster, aes(x = PerformanceScore, y = Value, fill = Variable)) + 
  geom_bar(stat = "identity", position = position_dodge(0.95), 
           color = design()$color$link_water, size = 0.7) + 
  geom_text(aes(label = Value), position = position_dodge(0.9), vjust = -0.7, size = 3) + 
  labs(
    title = "Work Life Balance of Employee",
    x = "Performance Score",
    y = "Variable"
  ) + 
  scale_fill_manual(values = c("Absences" = design()$color$gossip, 
                               "Emp_Satisfaction" = design()$color$amber, 
                               "Sp_Project" = design()$color$biscay)) + 
  theme_minimal() + 
  theme (
    plot.title = element_text(family = design()$font$title, 
                              colour = design()$color$biscay, size = 20),
    axis.title = element_text(family = design()$color$text, 
                              color = design()$color$biscay, size = 13),
    axis.text = element_text(family = design()$color$text, 
                             colour = design()$color$biscay, size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y = element_blank()
  )
```